### 2024-04-08 README.qmd
Add analyis

---
format: gfm
execute: 
  message: false
  warning: false
  echo: false
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

```{r}
#| echo: false
#| eval: false
# Package setup:
usethis::use_description()
usethis::use_package("sf")
usethis::use_package("stplanr")
usethis::use_package("ggplot2")
usethis::use_package("dplyr")
usethis::use_r("opsnap")
# MIT license:
usethis::use_mit_license("Leeds Institute for Transport Studies")

# Add CI
usethis::use_github_action_check_standard()
# Add pkgdown
usethis::use_pkgdown()
# Add pkgdown action
usethis::use_github_action("pkgdown")
# Add gh pages:
usethis::use_github_pages()
```

# Installation

Install the package from GitHub:

```r
remotes::install_github("ITSLeeds/opsnap")
```

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
remotes::install_github("ITSLeeds/opsnap")
```

```{r}
#| eval: false
#| echo: false
devtools::load_all()
```

# Opsnap data

The `opsnap` package provides a function to download and read in data from the West Yorkshire Police Operation Snap database. The data is available at the following URL: https://www.westyorkshire.police.uk/SaferRoadsSubmissions

Data for the following years are provided:

```{r}
#| echo: false
library(rvest)
query = ".file-link a"
url = "https://www.westyorkshire.police.uk/SaferRoadsSubmissions"
files_available = read_html(url) |>
  html_nodes(query) |>
  html_text()
links_available = read_html(url) |>
  html_nodes(query) |>
  html_attr("href")
urls = paste0("https://www.westyorkshire.police.uk", links_available)
file_names = basename(links_available)
tibble::tibble(file_names) |>
  knitr::kable()

# In preparation for reading-in all data:
date_str = format(Sys.Date(), "%Y-%m")
file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
```

```{r}
#| eval: false
#| echo: false
# Download them to raw_data/west-yorkshire:
dir.create("raw_data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
pbapply::pblapply(urls, function(u) {
  download.file(u, paste0("raw_data/west-yorkshire/", basename(u)))
})

# Test with 2nd url:
u = urls[2]
d2 = opsnap:::download_and_read(u)
d_all = purrr::map_df(urls, opsnap:::download_and_read)
dir.create("data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
# Date to nearest month:
write_csv(d_all, file_name)
```

The data is open acess and looks like this, with names cleaned up by the package:

```{r}
#| eval: false
#| echo: true
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
d = opsnap:::download_and_read(u)
names(d)
# Old names:
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"
# New names:
# [1] "mode"     "make"     "model"    "colour"   "offence"  "district" "disposal"
# [8] "date"     "location"
```

<!-- The data looks like this (first 3 rows shown): -->

```{r}
#| echo: false
d_all = read_csv(file_name)
d_all |>
  head(3) |>
  knitr::kable()
```

There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown below):

```{r}
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "all")
d = d_all |>
  opsnap:::filter_nas()
d_complete_monthly = d |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete")
d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
d_monthly |>
  ggplot() +
  geom_line(aes(month, n, colour = records)) +
  geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
       x = "Date",
       y = "Number of records per month")
```

As shown in the graph above, some records `r round(nrow(d) / nrow(d_all) * 100, 1)`% have missing values for locations (these records did not classify as an offence). Filtering these out leaves `r nrow(d)` complete records.

There are `r unique(d$location) |> length()` unique locations in the data, with the most common locations shown below:

```{r}
#| echo: false
d |>
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

# Geocoding

We provide a function to geocode the records:

```{r}
#| eval: false
#| echo: true
d_sample = d[1:5, ]
d_sf = opsnap:::op_geocode(d_sample)
mapview::mapview(d_sf)
```

```{r}
#| eval: false
#| echo: false
# Geocode all data:
d_geocoded = opsnap:::op_geocode(d)
d_geometries = d_geocoded |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
d_sf = sf::st_sf(
  d,
  geometry = d_geometries$geometry
)
west_yorkshire = pct::get_pct_zones("west-yorkshire")
sf::sf_use_s2(FALSE)
d_sf_wy = d_sf[west_yorkshire, ]
# Sanity check
d_sf_wy |>
  sample_n(1000) |>
  # mapview::mapview()
  plot()
# Save to file:
sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
```

# Analysis

Due to inaccuracy in the geocoding, we only know the locations of the records to within around 500m of each crash (although we can link to specific roads).
We'll present the geographic distribution of crashes using a 500m grid:

```{r}
d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
d_projected = sf::st_transform(d_sf, 27700)
library(terra)
raster_template = rast(ext(d_projected), res = 500)
raster_count = rasterize(d_projected, raster_template, fun = "length")
plot(raster_count)
```

The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again:

```{r}
raster_count_no_outlier = raster_count |>
  clamp(-Inf, 100)
plot(raster_count_no_outlier)
```

<!-- You can query the data downloaded with `opsnap` functions, e.g. as follows (results not shown): -->

```{r}
#| eval: false
#| echo: false
table(d$offence) |>
  sort()
```

<!-- Let's make a plot of the data: -->

```{r}
#| eval: false
#| echo: false
d |>
  # Reduce nchar of offence
  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
  group_by(offence) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d) / 50)|>
  ggplot() +
  geom_bar(aes(offence)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```